---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    results = "hold",
    comment = "#>"
)

# Package symbols used throughout, plus the shared mock HTTP router.
box::use(
    deribit[DeribitMarketData],
    connectcore[local_mock_api],
    ./tests/testthat/mock_router[.mock_routes]
)

# connectcore::local_mock_api installs the shared router (built from the deribit
# route table of FULLY SYNTHETIC fixtures) as httr2's global mock hook, so it
# intercepts every req_perform / req_perform_promise call and the document
# renders against deterministic data with no network.
local_mock_api(.mock_routes)
```

# deribit

**In plain terms:** Deribit is the world's biggest exchange for crypto options — contracts that are essentially bets on how much the price of Bitcoin or Ether will swing. Because so much of that activity happens in one place, it is the best window into what traders collectively expect the market to do next, including a single headline gauge, the DVOL index, that summarises how nervous or calm the options market is. This package pulls Deribit's public market data — the list of tradable contracts, live prices and order books, that volatility index, perpetual funding costs, and recent trades — and returns it as clean tables you can work with in R. It reads public data only: there is no trading and no login or API key required. It reports the exchange's numbers faithfully and leaves the interpretation to you.

## Technical overview

R API wrapper to the [Deribit](https://www.deribit.com) crypto options and derivatives exchange, supporting both synchronous and asynchronous (promise based) operations. Provides an R6 market-data client for the public JSON-RPC endpoints — instruments, ticker, order book, index price, DVOL volatility index, funding rates, trades, and option-chain book summaries — built on the shared [connectcore](https://github.com/dereckscompany/connectcore) transport base.

## What this is

Deribit is where the crypto **options** market lives — the deepest venue for BTC and ETH options, the perpetual and dated futures alongside them, and the **DVOL** volatility index that the whole desk watches as a regime dial. This package gives you Deribit's public market data as tidy `data.table`s: the instrument universe, live quotes and order books, DVOL candles, perpetual funding, recent trades, and the full option-chain snapshot.

It is a faithful, low-level wrapper: it speaks Deribit's JSON-RPC 2.0 wire format and returns clean tables, preserving the venue's own field names (only snake_cased), but it does not editorialise the data. **Phase 1 is public market data only** — there is no trading and no authenticated surface, and the public endpoints need no API key.

## Design philosophy

- **`data.table` everywhere, no list columns.** Every method returns one flat `data.table`; measurement columns (and any field legitimately absent for a given instrument kind — an option's greeks, a perpetual's funding, a null bid) are typed nullable, structural columns strict.
- **Sync and async.** Every request-making surface works in both modes. `async = TRUE` returns a [promise][promises::promise]; otherwise the table is returned directly. There is a single sync/async branch point (inherited from `connectcore`).
- **Faithful field names.** Venue-native names are preserved and only snake_cased; nested objects (a ticker's `stats`/`greeks`) are flattened with the parent field as a prefix (`stats_high`, `greeks_delta`).
- **Typed errors.** Every failure is a classed condition (`deribit_api_error`, `deribit_validation_error`) carrying structured fields — you branch on the type, never grep the message.

## Installation

This project uses [`renv`](https://rstudio.github.io/renv/). Add `deribit` to your lockfile and restore:

```{r install, eval = FALSE}
renv::install("dereckscompany/deribit")
# or, without renv:
# remotes::install_github("dereckscompany/deribit")
```

## The instrument universe

Every method hangs off one client. The public endpoints need no key, so construction takes no credential.

```{r instruments}
md <- DeribitMarketData$new()

options <- md$get_instruments("BTC", kind = "option")
options[, .(instrument_name, option_type, strike, expiration_datetime)]
```

`get_currencies()` lists the settlement currencies, and `kind` filters the universe (`"future"`, `"option"`, `"spot"`, `"future_combo"`, `"option_combo"`).

## Live quotes

The ticker combines the 24h stats, the top of book, and — for options — the greeks and implied volatility, flattened into one row:

```{r ticker}
perp <- md$get_ticker("BTC-PERPETUAL")
perp[, .(instrument_name, mark_price, index_price, current_funding, funding_8h)]
```

The order book comes back long — one row per price level, bids then asks:

```{r order-book}
md$get_order_book("BTC-PERPETUAL", depth = 3)
```

## The DVOL volatility index

`get_volatility_index_data()` returns the DVOL candles over a window — the regime dial. Timestamps are `lubridate` POSIXct; a single call returns up to ~1000 candles, and a longer range pages by moving `end_timestamp` back:

```{r dvol}
dvol <- md$get_volatility_index_data(
    "BTC",
    start_timestamp = lubridate::ymd_hms("2023-11-01 00:00:00", tz = "UTC"),
    end_timestamp = lubridate::ymd_hms("2023-11-08 00:00:00", tz = "UTC"),
    resolution = "3600"
)
dvol
```

## Perpetual funding

```{r funding}
funding <- md$get_funding_rate_history(
    "BTC-PERPETUAL",
    start_timestamp = lubridate::ymd_hms("2023-11-01 00:00:00", tz = "UTC"),
    end_timestamp = lubridate::ymd_hms("2023-11-08 00:00:00", tz = "UTC")
)
funding
```

## The option chain

`get_book_summary_by_currency(currency, kind = "option")` is the whole chain in one call — mark IV, open interest, volume, and the (null-able) bid/ask/mid for every live option. `get_option_chain()` is the convenience wrapper:

```{r chain}
chain <- md$get_option_chain("BTC")
chain[, .(instrument_name, mark_iv, bid_price, ask_price, open_interest, volume)]
```

## Asynchronous usage

Set `async = TRUE` and consume the promise with `coro::async` / `await`, driving the event loop with `later`:

```{r async, eval = FALSE}
box::use(coro, later)

md_async <- DeribitMarketData$new(async = TRUE)

main <- coro::async(function() {
    dvol <- await(md_async$get_volatility_index_data(
        "BTC",
        start_timestamp = lubridate::now("UTC") - lubridate::days(7),
        end_timestamp = lubridate::now("UTC"),
        resolution = "3600"
    ))
    chain <- await(md_async$get_option_chain("BTC"))
    print(list(dvol = dvol, chain = chain))
})

main()
while (!later::loop_empty()) later::run_now()
```

## Error handling

A bad argument is caught before any request as a `deribit_validation_error`; a Deribit JSON-RPC error surfaces as a typed `deribit_api_error` carrying the HTTP status, the JSON-RPC code, and the offending parameter:

```{r errors}
result <- tryCatch(
    md$get_volatility_index_data("BTC", lubridate::now("UTC"), lubridate::now("UTC"), resolution = "99"),
    deribit_validation_error = function(e) paste("caught:", conditionMessage(e))
)
result
```
